---
title: "Developing Data Products Final Project"
author: "Andrew Delos Santos"
date: "9/24/2019"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Project Summary

This R Markdown presentation is made interactive using Shiny. The app gives a 
blank map of Italy and helps you see the different regions within Italy. When
you select from the drop down menu, it allows you to pick a region which will
then be highlighted in red. The output below shows the first 6 regions in Italy.

```{r echo = FALSE}
library(shiny)
library(maps)
library(mapdata)
library(ggplot2)

italy_states <- map_data("italy")
head(italy_states)
```

## How it Operates
-The app is operated through shiny with includes two source code files

-The first one is the **server.R** source code and the other is the **ui.R**
 source code
 
-The app uses both the libraries maps and mapdata to produce the Italy map

## Instructions
-From the menu, pick a region in Italy

-The chosen region will appear red

## Links to App and Code on Github
https://github.com/Andrew20663/DLSPortfolio
https://andrew20663.shinyapps.io/developing_data_products/

