---
title: "Developing Data Products Final Project"
author: "Andrew Delos Santos"
date: "9/18/2019"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This final project for the Developing Data Products course is based off of the Regression Models Final Project that was done earlier in the specialization. Previously, we analyzed the relationship between a certain set of variables and the miles per gallon. We aimed to answer two questions:  
    
    1. "Is an automatic or manual transmission better for MPG?"
    2. "Quantify the MPG difference between automatic and 
        manual transmissions." 

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0, max = 1, value = 1, step = 0.1)
)

renderPlot({
multivariable_model <- lm(mtcars$mpg ~ mtcars[["am"]] + mtcars[["cyl"]] + 
mtcars[["disp"]] + mtcars[["hp"]] + mtcars[["wt"]], mtcars)
r <- resid(multivariable_model)
hist(r, probability = TRUE)
  
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



