---
title: "Regression Models Project"
output:
  pdf_document: 
  fontsize: 5 pt
  geometry: margin = 0.5 in
---
### Executive Summary

In this particular report, I looked at a data set of collection of cars and was 
interested in exploring the relationships between a set of variables and miles 
per gallon (the outcome). The goal was to answer two particular questions: 
  1) "Is an automatic or manual transmission better for MPG?"
  2) "Quantify the MPG difference between automatic and 
      manual transmissions." 
These two questions were answered in which we saw that automatic transmission 
is worse for MPG and that the manual is better for mpg in the given boxplot. 
From the regression analysis, we were able to quantify these values in which
automatic had a slope of 17.14737 and that manual had a slope of 24.39231. 
The difference between these two values is 7.24494. The interpretation of each 
slope is that for there is that for every 17.14737 miles gone, 1 gallon is used.
Also, for every 24.39231 miles gone, one gallon is used. However, the R^2 value
represents that only 35.98% of the variation in y (mpg) can be explained by x
(transmission). Therefore, I created more a multivariable model and included 
this predictor since it was indeed significant at a .05 alpha significance level.
From the multivariable model, the model was also significant and the R^2 value 
much larger where adjusted R^2 was 0.8344. 83.44% of the variation in mpg could
be explained by the tranmsission, cylinder, horsepower and weight of the car which
were significant at the .05 level. Also, the AIC and BIC values were lower than 
that of the single variable model which is a better model. The residuals were also
approximately normal through plots and the Shapiro test. So overall, both questions
were answered and the multivariable model was accurate.

### Exploratory Data Analysis
```{r echo = TRUE}
#Get some information regarding the data set mtcars
?mtcars
#Load the mtcars data set
data(mtcars)
#See the structure of the data along with each variable
#and their particular class
str(mtcars)
#See the first 6 values
head(mtcars)
#See the names of the variables for the data set
names(mtcars)
#Convert transmission, engine shape, carbeurator, cylinder
#and gear to factor variables
mtcars[["am"]] <- factor(mtcars[["am"]], levels = c(0,1), labels = c("Auto", "Manual"))
mtcars[["vs"]] <- factor(mtcars[["vs"]])
mtcars[["carb"]] <- factor(mtcars[["carb"]])
mtcars[["cyl"]] <- factor(mtcars[["cyl"]])
mtcars[["gear"]] <- factor(mtcars[["gear"]])
#Exploratory plots are in the appendix
```

### Regression Analysis
```{r echo=TRUE}
#Subset the data of mtcars outcome mpg as a linear model
#or function of transmission type and calculate the mean
#for the subset
aggregate(mtcars[["mpg"]] ~ mtcars[["am"]], mtcars, mean)
#Create two variables to set boolean equal to character string of Auto and 
#manual
a <- mtcars[mtcars[["am"]] == "Auto",]
m <- mtcars[mtcars[["am"]] == "Manual",]
#Conduct t-test both subsetted data
t.test(a[["mpg"]], m[["mpg"]])
auto_manual_t_test <- t.test(a[["mpg"]], m[["mpg"]])
#Create confidence interval
auto_manual_t_test[["conf.int"]]
#See the validity of a linear model with just mpg and transmission type 
linear_model_1 <- lm(mtcars$mpg ~ mtcars[["am"]], mtcars)
#Calculate the summary and AIC and BIC values
summary(linear_model_1)
AIC(linear_model_1)
BIC(linear_model_1)
#Values indicate a positive outcome, add more predictors to the model from the
#pairs plot in the appendix 
multivariable_model <- lm(mtcars$mpg ~ mtcars[["am"]] + mtcars[["cyl"]] + 
mtcars[["disp"]] + mtcars[["hp"]] + mtcars[["wt"]], mtcars)
#See analysis of variance summary on both models
anova(linear_model_1, multivariable_model)
#See just the summary for the multi-variable model and see the AIC and BIC values
summary(multivariable_model)
AIC(multivariable_model)
BIC(multivariable_model)
#Calculate the residuals of the multi-variable model
r <- resid(multivariable_model)
#Regression plots are in the appendix
```

### Appendix
```{r echo = TRUE}
#Exploratory Analysis Plots

#Create boxplot to see quartiles and if there are any outliers
library(ggplot2)
ggplot(mtcars, aes(mtcars[["am"]], mtcars[["mpg"]], fill = mtcars[["am"]])) +
geom_boxplot() + facet_wrap(. ~ mtcars[["am"]]) + xlab("Transmission Type") +
ylab("MPG") + guides(fill = guide_legend(title = "Transmission Type"))
#Create plot to see which variables seem to have a correlation
pairs(mtcars$mpg ~ ., mtcars)

#Regression Analysis Plots
#Analyze the residuals and conduct tests for normality
par(mfrow = c(3,1))
hist(r, probability = TRUE)
lines(density(r), col = "red")
qqnorm(r)
qqline(r, col = "red")
var(r)
shapiro.test(r)
plot(r)
```