---
title: "Statistical Inference: Siumulation Exercise"
output: pdf_document
---
###Part 1: Simulation Exercise

Synopsis: For this particular project I invesigated the exponential distribution
in R and compared it with the Central Limit Theorem. The exponential distribution
has a rate parameter of lambda where mean is 1/lambda. I ran a simulation on 40 
exponentials. 

###Simulation of Mean of Exponential Random Variable
```{r echo = TRUE}
#Set given parameters for simulation of the exponential random variable
#Use rexp function with 1000 simulations of 40 exponentials
set.seed(10)
simulation <- data.frame(ncol = 2, nrow = 1000)
names(simulation) <- c("simulation", "mean")
#Create for loop to build the data frame for 1-1000 where column 1 is filled
#with the number of the simulation and the 2nd column is the mean of the 
#random exponential distribution with given parameters n=40 and lambda=0.2
for(i in 1:1000) {
  simulation[i,1] <- i
  simulation[i,2] <- mean(rexp(40,0.2))
  }
```


###Sample Mean vs. Theoretical Mean:
From Central Limit Theorem, an exponential follows distribution with mean 1/lam.
The mean is 5 when you calculate it. 
```{r echo=TRUE}
#Calculate both theoretical and sample mean
theoretical_mean <- 1/0.2
simulated_mean <- mean(simulation[["mean"]])
print(theoretical_mean)
print(simulated_mean)
```

###Histogram of the Simulation of Means vs. Theoretical Mean Line
```{r echo = TRUE}
library(ggplot2)
#Create histogram plot with density on y-axis and mean on x-axis
#Red vertical line shows theoretical mean while green shows mean of simulated
#exponential random variables
ggplot(simulation, aes(mean, ..density..)) + geom_histogram(col = "blue") +
xlab("Simulated mean") + ylab("Density") + ggtitle("Simulated Mean of 
Exponential Random Variable") + geom_vline(xintercept=5, color = "red") +
geom_vline(xintercept = 5.04506, color = "green") + 
theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
                                                    
```


###Sample Variance vs. Theoretical Variance
```{r echo =TRUE}
set.seed(10)
#Set given parameters for simulation of the exponential random variable
#Use rexp function with 1000 simulations of 40 exponentials
set.seed(10)
simulation_var <- data.frame(ncol = 2, nrow = 1000)
names(simulation_var) <- c("simulation", "variance")
#Same as in mean step except this time for the variance 
for(i in 1:1000) {
  simulation_var[i,1] <- i
  simulation_var[i,2] <- mean(rexp(40,0.2))
  }
```


```{r echo = TRUE}
#Calculate both theoretical and simulated means variance
theoretical_variance <- 1/(0.2^2)/40
simulated_variance <- (sd(simulation_var[["variance"]]))^2
print(theoretical_variance)
print(simulated_variance)
```

###Histogram of Sample Variance 
```{r echo = TRUE}
library(ggplot2)
#Plot histogram
ggplot(simulation_var, aes(variance, ..density..)) + geom_histogram(col = "blue") +
xlab("Simulated Mean") + ylab("Density") + ggtitle("Variance of Simulated Means
vs. Theoretical for Exponential Random Variable") +
theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
```

###Check if Exponential is approximately Normal
```{r echo=TRUE}
library(ggplot2)
#Create Normal Distribution curve to be put around the histogram 
x_axis_norm <- seq(min(simulation[["mean"]]), max(simulation[["mean"]]), 
length.out = 1000)
y_axis_norm <- dnorm(x_axis_norm, 1/0.2, (1/.2)/sqrt(40), log = FALSE)

#Plot the historgram
ggplot(simulation, aes(mean, ..density..)) + geom_histogram(col = "blue") +
xlab("Simulated mean") + ylab("Density") + ggtitle("Simulated Mean of 
Exponential Random Variable") + geom_vline(xintercept=5, color = "red") +
geom_vline(xintercept = 5.04506, color = "green") + 
geom_line(aes(x_axis_norm,y_axis_norm)) + 
theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
```
###Results: 

Overall, in the above figures, we see that the 40 exponential random
variables simulation does follow a normal distribution with 1000 simulations.
In the first plot, we see that the theoretical mean is 5 represented by the red 
line and the mean ofthe simulated random variables is 5.04506 shown by the green
line. The final distribution showing that the data is approximately normal.




